Mechanistic Bayesian Forecasts of COVID19

### 2020-07-30 paper.tex
update

\begin{document}
\section*{Abstract}

@@

Compartmental models have also been adopted into a Bayesian framework before, including both stochastic disease dynamics and deterministic dynamics \cite{hotta2010bayesian}\cite{dukic2012tracking}. In fact, non-parametric transmissibility was included in a Bayesian SEIR model to study Ebola by Frasso and Lambert \cite{frasso2016bayesian}. Time-varying transmissibility has also been studied in the frequentist setting \cite{smirnova2019forecasting}. With the outbreak of COVID-19, accounting for testing has become a critical element in using an SEIR model. Lopez et al. used a fixed detection probability to model undiagnosed individuals in Spain and Italy \cite{lopez2020modified}. Many efforts have been made to use SEIR models in forecasting COVID-19 \cite{giordano2020modelling}\cite{yang2020modified} \cite{bertozzi2020challenges}\cite{prem2020effect}.

@@

\begin{figure}
  \centering

@@

\caption{Example fit and forecast for states with over 50 incident deaths.}
\label{fig:fit_and_forecast_results}
\end{figure}

\begin{figure}
  \centering
     \begin{subfigure}{.5\textwidth}
  \centering
    \includegraphics[scale=.1]{beta_t_plot.png}
    \caption{Time varying-beta.}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
    \includegraphics[scale=.1]{detection_plot.png}
    \caption{Time-varying detection probability}
\end{subfigure} 
 \caption{  A) Time varying beta parameter for four example states. We can clearly see that the random walk is able to non-parametrically account for the non-pharamceutical interventions. B) Time varying detection probability for four example states. The time varying detection random walk is able account for the increase in testing. }
\label{fig:model_details}
\end{figure}

@@

We can see from Figure \ref{fig:fit_and_forecast_results} that MechBayes version 3 is able to accurately model the observed data in the three example states (NY,CA,FL) as well as the US. The model is able to adapt to highly variable incident death reporting, variable transmission rates, and overall heterogeneity of incidence curves. The model is also able capture the uncertainty of the differential equation parameters well enough to produce well calibrated prediction intervals. Figure \ref{fig:fit_and_forecast_results}  shows intervals at the 50\%,80\% and 95\% levels. However, we can also see that in some states, such as California and Florida, the model is biased high, with all observations outside of the 95\% prediction interval falling below. Figure \ref{fig:fit_and_forecast_results}  also shows 4 weeks of daily forecasts, along with the daily observed incidence for 1 week out. We can see that the predictions are tracking the data even under the weekly reporting cycles.

@@

\item \textbf{MechBayes is biased high}. We can see from Figure \ref{fig:bias} that MechBayes was biased high with respect to the median forecast. We suspect this is due to the inherent exponential growth of an SEIR model when the number of susceptibles is small. Since overall prevalence in any state is well below the number needed to reach herd immunity, when the model sees an increase in cases it treats this as exponential growth, which translates into exponential growth in deaths. However, in the recent forecasts the bias is consistently lower than the COVID-HUB baseline model. Note the extreme low bias in the first submission is due to the omission of the U.S. in the submission. \textbf{is this true} \textbf{something is telling me this is also related to the exponential random walk}

@@

@@

\begin{figure}
\includegraphics[scale=.2]{bias_by_timezero.png}

\caption{Bias of MechBayes and COVID-Baseline as a function of time. }
\label{fig:bias}
\end{figure}

>>>>

\begin{document}

\section*{Abstract}

@@

Compartmental models have also been adopted into a Bayesian framework before, including both stochastic disease dynamics and deterministic dynamics \cite{hotta2010bayesian}\cite{dukic2012tracking}. In fact, non-parametric transmissibility was included in a Bayesian SEIR model to study Ebola by Frasso and Lambert \cite{frasso2016bayesian}. Time-varying transmissibility has also been studied in the frequentist setting \cite{smirnova2019forecasting}. With the outbreak of COVID-19, accounting for testing has become a critical element in using an SEIR model. Lopez et al. used a fixed detection probability to model undiagnosed individuals in Spain and Italy \cite{lopez2020modified}. Many efforts have been made to use SEIR models in forecasting COVID-19 \cite{giordano2020modelling}\cite{yang2020modified} \cite{bertozzi2020challenges}\cite{prem2020effect}\cite{flaxman2020estimating}. Perhaps the most similar model was put forth by Pei et al. \cite{pei2020differential}. This model was an extension to the SEIR model with time varying transmission and detection probability. However, their model fixed detection probability across time and was fit using Kalman Filter techniques instead of Bayesian HMC. They also used the model mostly for counterfactual scenario projections instead of forecasting. Another relatively similar model was put forth by Abbot et al. \cite{abbott2020estimating}. This model does use a time-varying transmission rate and detection rate, but does not use a differential equation model as the core component, but rather parameterizes new cases as a function of the time varying reproduction number.

@@

\begin{figure}

@@

\caption{Example fit and forecast for states with over 50 incident deaths. Grey bands represent 95\% prediction intervals. Blue line represents median forecast. MechBayes is able to produce well calibrated fits to the data, even under extreme reporting variation such as in North Carolina. It is also able to handle extreme outliers, such as in Illinois or New Jersey. \textbf{Is this totally true or should we mention manual adjustments.}}
\label{fig:fit_and_forecast_results}
\end{figure}

\begin{figure}
  \centering
     \begin{subfigure}{1\textwidth}
  \centering
    \includegraphics[scale=.2]{beta_t_plot.png}
    \caption{Time varying-beta.}
\end{subfigure}

\begin{subfigure}{1\textwidth}
  \centering
    \includegraphics[scale=.2]{detection_plot.png}
    \caption{Time-varying detection probability}
\end{subfigure} 
 \caption{  A) Time varying beta parameter for four all geographies with a day exceeding 50 incident deaths. We can clearly see that the random walk is able to non-parametrically account for the non-pharmaceutical interventions. B) Time varying detection probability for four example states. The time varying detection random walk is able account for the increase in testing. }
\label{fig:model_details}
\end{figure}

@@

We can see from Figure \ref{fig:fit_and_forecast_results} that MechBayes version 3 is able to accurately model the observed data. The model is able to adapt to highly variable incident death reporting, variable transmission rates, and overall heterogeneity of incidence curves. The model is also able capture the uncertainty of the differential equation parameters well enough to produce well calibrated prediction intervals. Figure \ref{fig:fit_and_forecast_results}  shows prediction intervals at the 95\% level. However, we can also see that in some states, such as California and Florida, the model is biased high, with all observations outside of the 95\% prediction interval falling below. Figure \ref{fig:fit_and_forecast_results}  also shows 4 weeks of daily forecasts, along with the daily observed incidence for 1 week out. We can see that the predictions are tracking the data even under the weekly reporting cycles.

@@

\item \textbf{MechBayes is biased high}. We can see from Figure \ref{fig:results_discussion} that MechBayes was biased high with respect to the median forecast. We suspect this is due to the inherent exponential growth of an SEIR model when the number of susceptibles is small. Since overall prevalence in any state is well below the number needed to reach herd immunity, when the model sees an increase in cases it treats this as exponential growth, which translates into exponential growth in deaths. However, in the recent forecasts the bias is consistently lower than the COVID-HUB baseline model. Note the extreme low bias in the first submission is due to the omission of the U.S. in the submission. \textbf{is this true} \textbf{something is telling me this is also related to the exponential random walk}

@@

\item \textbf{MechBayes Full outperforms the other versions on probabilistic scoring measures}. As can be seen by Figure \ref{fig:results_discussion}, MechBayes Full achieves the best log-score \textbf{sub for WIS } out the three nested models. This suggested that our extensions to the basic SEIR model not only improve point-forecasts, but also calibration.

@@

\begin{figure}
\begin{subfigure}{.5\textwidth}
\includegraphics[scale=.15]{bias_by_timezero.png}
\subcaption{Bias of MechBayes and COVID-Baseline as a function of time. }
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
\includegraphics[scale=.15]{ablation_calibration.png}
\subcaption{Comparison of nested models using the probabilistic scoring mechanism}
\end{subfigure}
\caption{A) Bias of MechBayes over time. B) Probabilistic Calibration}
\label{fig:results_discussion}
\end{figure}
